---
title: "Discussion12"
author: "Mary Anna Kivenson"
date: "November 16, 2019"
output:
  rmarkdown::html_document:
    theme: cosmo 
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)
library(readxl)
```


# Multiple Linear Regression {.tabset}

## Dataset

The dataset I will be using is from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Concrete+Compressive+Strength), containing concrete strength factor data. 

The dataset has the following factors:

* Cement (component 1) -- quantitative -- kg in a m3 mixture -- Input Variable
* Blast Furnace Slag (component 2) -- quantitative -- kg in a m3 mixture -- Input Variable
* Fly Ash (component 3) -- quantitative -- kg in a m3 mixture -- Input Variable
* Water (component 4) -- quantitative -- kg in a m3 mixture -- Input Variable
* Superplasticizer (component 5) -- quantitative -- kg in a m3 mixture -- Input Variable
* Coarse Aggregate (component 6) -- quantitative -- kg in a m3 mixture -- Input Variable
* Fine Aggregate (component 7) -- quantitative -- kg in a m3 mixture -- Input Variable
* Age -- quantitative -- Day (1~365) -- Input Variable

And the following dependent variable:

* Concrete compressive strength -- quantitative -- MPa -- Output Variable

We preview the data below.

```{r}
concrete <- read.csv("https://raw.githubusercontent.com/mkivenson/Computational-Mathematics/master/Discussions/Concrete%20Strength%20Dataset.csv", header = TRUE)

head(concrete)
```


## Exploring the Data

#### Distribution of Cemete Strength

The distribution of cement strength is normal and centered at 35.82. 

```{r}
ggplot(concrete, aes(Strength)) + 
  geom_histogram(bins = 25) + 
  ggtitle("Distribution of Cemete Strength")

summary(concrete$Strength)
```


#### Correlation Plot

The following chart looks at correlation between different factors and cement strength.

```{r}
library(corrplot)
corrplot(cor(concrete), 
         type = "upper", 
         order = "hclust")
```


## Regression

For multiple linear regression, we use Fly Ash, Superplasticizer, Blast Furnace Slag, Water, Age, and Cement as factors. The p-value for Fine Aggregate and Course Aggregate is > 0.05. 

```{r}
lm_concrete <- lm(Strength ~ +
                    #Fine.Aggregate + 
                    Fly.Ash + 
                    Superplasticizer + 
                    Blast.Furnace.Slag + 
                    Water + 
                    Age + 
                    #Course.Aggregate + 
                    Cement, 
                  data = concrete)
summary(lm_concrete)
```

#### Multiple Regression Line

$$ y = 29 + 0.069*\hat{Fly.Ash } + 0.24*\hat{Superplasticizer} + 0.086*\hat{Blast.Furnace.Slag} 
-0.218*\hat{Water} + 0.113*\hat{Age} + 0.105*\hat{Cement}$$


## Residual Analysis


#### QQ Plot

```{r}
qqnorm(lm_concrete$residuals)
qqline(lm_concrete$residuals)
```


#### Histogram of Residuals

The histogram of residuals is normally distributed.

```{r}
hist(lm_concrete$residuals)
```

#### Residual Plot

The residual plot shows constant variability, but seems to have a linear pattern. 

```{r}
plot(lm_concrete$residuals ~ concrete$Strength)
abline(h = 0, lty = 3)
```











