---
title: "Assignment12"
author: "Mary Anna Kivenson"
date: "November 17, 2019"
output:
  rmarkdown::html_document:
    theme: cosmo 
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Assignment 12 {.tabset}

## Dataset

The attached who.csv dataset contains real-world data from 2008. The variables included follow.

* Country: name of the country

* LifeExp: average life expectancy for the country in years

* InfantSurvival: proportion of those surviving to one year or more

* Under5Survival: proportion of those surviving to five years or more

* TBFree: proportion of the population without TB.

* PropMD: proportion of the population who are MDs

* PropRN: proportion of the population who are RNs

* PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate

* GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate

* TotExp: sum of personal and government expenditures.

```{r}
who <- read.csv("https://raw.githubusercontent.com/mkivenson/Computational-Mathematics/master/Assignments/Assignment%2012/who.csv")

head(who)
```


## Step 1

Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the
variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss
whether the assumptions of simple linear regression met.

#### Scatterplot

```{r}
ggplot(who, aes(TotExp, LifeExp)) + geom_point() + 
  ggtitle("Scatterplot of LifeExp~TotExp") 
```


#### Linear regression

```{r}
attach(who)
lm_who <- lm(LifeExp~TotExp)
summary(lm_who)
```


#### Assumptions

* Linear relationship - Not met, the scatterplot is not linear.
* Normally distributed Residuals - Not met, the residuals are skewed.

```{r}
ggplot(lm_who, aes(x=lm_who$residuals)) + 
  geom_histogram(binwidth=1) +
  ggtitle("Distribution of Residuals")
```

* Independence of the observations - population health a country must be independent of population health of other countries (assume true for this exercise?)
* Homoscedasticity - Variance appears constant

#### Results

* The F-statistic of 65.26 is the overall significance of the regression model. 
* The R^2 of 0.2577 means that the fit of the model is subpar
* The Standard Error of 9.371 is the amount which the estimated life expectance varies from the actual value.
* The p-value of 7.71e-14 indicates the relationship between LifeExp and TotExp is significant

## Step 2

Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06
power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r
re-run the simple regression model using the transformed variables. Provide and interpret the F
statistics, R^2, standard error, and p-values. Which model is "better?"

#### Transformations

```{r}
who$LifeExp <- who$LifeExp^4.6
who$TotExp <- who$TotExp^.06
```


#### Linear regression

```{r}
ggplot(who, aes(TotExp, LifeExp)) + geom_point() + 
  ggtitle("Scatterplot of LifeExp~TotExp") 
```

#### Scatterplot

```{r}
attach(who)
lm_who <- lm(LifeExp~TotExp)
summary(lm_who)
```


#### Assumptions


* Linear relationship - Met, the adjusted relationship is moderately strong, positive, and linear.
* Normally distributed Residuals - Met, the residual histogram is normal.


```{r}
ggplot(lm_who, aes(x=lm_who$residuals)) + 
    geom_histogram(bins = 20) +
  ggtitle("Distribution of Residuals")
```

* Independence of the observations - population health a country must be independent of population health of other countries (assume true for this exercise?)
* Homoscedasticity - Variance in the scatterplot appears constant


#### Results
* The F-statistic of 507.7 is the overall significance of the regression model. 
* The R^2 of 0.7298 means that the model fits the data well
* The Standard Error of 90490000 is the amount which the estimated life expectance^4.6 varies from the actual value.
* The p-value of <2e-16 indicates the relationship between LifeExp and TotExp is significant

## Step 3

Using the results from 3, forecast life expectancy when TotExp^.06 = 1.5. Then forecast life expectancy when TotExp^.06=2.5. 

#### Model
$$LifeExp^{4.6} = -736527910 + 620060216 * TotExp^.06$$

```{r}
exp_model <- function(x){
  y = -736527910 + (620060216 * x)
  return(y)
}
```


#### TotExp^.06 = 1.5

```{r}
exp_model(1.5)
```


#### TotExp^.06 = 2.5

```{r}
exp_model(2.5)
```

## Step 4

Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp


#### Multiple regression

```{r}
attach(who)
mult_lm_who <- lm(LifeExp ~ + 
               PropMD + 
               TotExp +
               (PropMD * TotExp))
summary(mult_lm_who)
```


#### Results
* The F-statistic of 180.3 is the overall significance of the regression model. 
* The R^2 of 0.7441 means that the model fits the data well
* The Standard Error of 88520000 is the amount which the estimated life expectance^4.6 varies from the actual value.
* The p-value of <2e-16 indicates the linear model is significant

## Step 5

Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

#### Prediction

```{r}
mult_exp_model <- function(a,b){
  y = -724400000 + 
    (47270000000 * a) + 
    (604800000 * b) + 
    (-21210000000 * a * b)
  return(y)
}
(mult_exp_model(0.03, 14))^(1/4.6)
```

This result seems reasonable because we used the adjusted exponential model. I will check results with the original dataset as well.


## Step 5 (No Adj)

Here, I re-did the multiple linear regression with the original dataset (without adjusting LifeExp and TotExp).

```{r}
who <- read.csv("https://raw.githubusercontent.com/mkivenson/Computational-Mathematics/master/Assignments/Assignment%2012/who.csv")
```

#### Multiple regression

```{r}
attach(who)
mult_lm_who <- lm(LifeExp ~ + 
               PropMD + 
               TotExp +
               (PropMD * TotExp))
summary(mult_lm_who)
```


#### Results
This model is much less accurate than the adjusted version, with an R^2 of 0.3574.


#### Prediction

The prediction of 107.68 years does not seem accurate, suggesting that this model is not accurate. 

```{r}
mult_exp_model <- function(a,b){
  y = 62.77 + 
    (1497 * a) + 
    (.0000723300 * b) + 
    (-.006026 * a * b)
  return(y)
}
mult_exp_model(0.03, 14)
```
