---
title: "Computational Mathematics - Discussion 13"
author: "Mary Anna Kivenson"
date: "November 21, 2019"
output:
  rmarkdown::html_document:
    
    theme: cosmo 
    highlight: zenburn
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(ggfortify)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


# Calculus

## Question 3 {.tabset}


Water flows onto a flat surface at a rate of 5 cm^3/s forming a
circular puddle 10mm deep. How fast is the radius growing
when the radius is:

volume of puddle = area of circle x depth
depth = 10mm = 1cm

$$\frac{dV}{dt} = \pi 2r \frac{dr}{dt} h$$
$$\frac{dr}{dt} = \frac{\frac{dV}{dt}}{2 \pi rh}$$

### 1 cm

```{r}
cyl_radius_change <- function(dv, radius, height){
  dr = dv/(2 * pi * radius * height)
  return(dr)
}
print('The radius change over time (cm/s) when the radius is 1cm is:')
cyl_radius_change(5, 1, 1)
```


### 10 cm

```{r}
print('The radius change over time (cm/s) when the radius is 10cm is:')
cyl_radius_change(5, 10, 1)
```


### 100 cm

```{r}
print('The radius change over time (cm/s) when the radius is 100cm is:')
cyl_radius_change(5, 100, 1)
```

## Question 4 {.tabset}


A circular balloon is inflated with air flowing at a rate of
10 cm^3/s. How fast is the radius of the balloon increasing
when the radius is:

$$V = \frac{4}{3} \pi r^3$$
$$\frac{dV}{dt} = 4 \pi r ^2 \frac{dr}{dt}$$
$$\frac{dr}{dt} = \frac{\frac{dV}{dt}}{4 \pi r^2}$$

### 1 cm

```{r}
sph_radius_change <- function(dv, radius){
  dr = dv/(4 * pi * (radius^2))
  return(dr)
}
print('The radius change over time (cm/s) when the radius is 1cm is:')
sph_radius_change(10, 1)
```

### 10 cm

```{r}
print('The radius change over time (cm/s) when the radius is 10cm is:')
sph_radius_change(10, 10)
```


### 100 cm

```{r}
print('The radius change over time (cm/s) when the radius is 100cm is:')
sph_radius_change(10, 100)
```