---
title: "Computational Mathematics - Assignment 2"
author: "Mary Anna Kivenson"
date: "September 7, 2019"
output:
  rmarkdown::html_document:
    
    theme: simplex
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Problem Set 1


#### Part 1

Show that: 
$$A^TA \neq AA^T $$

**Proof:** 

One of the basic laws of matrices is:

$$AB \neq BA $$


This law shows that matrix multiplication is not commutative, therefore $$A^TA \neq AA^T $$

**Demonstation:**

Generate a random matrix A:

```{r q1-1a}
matrix_A = matrix(sample(1:20, 9, replace = TRUE), nrow = 3, ncol = 3)
print(matrix_A)
print(t(matrix_A))
```


$$A^TA:$$

```{r q1-1b}
print(t(matrix_A) %*% matrix_A)
```


$$AA^T: $$

```{r q1-1c}
print(matrix_A %*% t(matrix_A))
```


The example above shows that generally, the products of matrix multiplication are not equal when multiplication order is inversed.



#### Part 2
For a special type of square matrix A, we get: 
$$A^TA = AA^T $$ 
Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).



This is true when the transpose of A is equal to A.

**Demonstration:**

```{r q1-2a}
matrix_B = matrix(c(1,0,1,0,1,0,1,0,1), nrow = 3, ncol = 3)
matrix_B
```


```{r q1-2b}
matrix_B %*% t(matrix_B) == t(matrix_B) %*% matrix_B
```



### Problem Set 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer.

To solve this question, I used an algorithm from [this site](https://www.geeksforgeeks.org/doolittle-algorithm-lu-decomposition/?fbclid=IwAR3tICsTPRsSi579Ct_q2ZeV-YiebwC3dJ7mFUBx-ml11AX2qzi0Y-Nu2lg).


```{r 2a}
lu <- function(matrix){

  n = nrow(A)
  
  lower = matrix((c(1:n)*0), nrow=n, ncol=n)
  upper = matrix((c(1:n)*0), nrow=n, ncol=n)
  
  for(i in (1:n)) {
    for(k in (i:n)) {
      sum = 0
      for(j in (1:i)) {
        sum = sum + (lower[i,j] * upper[j, k])
      }
      upper[i, k] = A[i, k] - sum
    }
    for(k in (i:n)) {
      if(i == k) {
        lower[i, i] = 1
      } else {
        sum = 0
        for(j in (1:i)) {
          sum = sum + (lower[k, j] * upper[j, i])
        }
        lower[k, i] = (A[k, i] - sum) / upper[i, i]
      }
    }
  }
  
  return(list(lower, upper))
  }
```

Generate a random 3x3 matrix:

```{r 2b}
A <- matrix(sample(1:20, 9, replace = TRUE), nrow = 3, ncol = 3)
print(A)
```

The function returns the following lower form matrix:

```{r 2c}
#lower:
lu(A)[[1]]
```

The function returns the following upper form matrix:

```{r 2d}
#upper:
lu(A)[[2]]
```

Test that the product of the lower and upper form matrices is equal to the original matrix:

```{r 2e}
#test: 
lu(A)[[1]] %*% lu(A)[[2]] == A
```




