---
title: "Computational Mathematics - Assignment 10"
author: "Mary Anna Kivenson"
date: "November 2, 2019"
output:
  rmarkdown::html_document:
    
    theme: cosmo 
    highlight: zenburn
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(ggfortify)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


# Probability - HW 10

## Question 1 {.tabset}

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him.

If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if

### Part A

(a) he bets 1 dollar each time (timid strategy).


#### Theorem

$$ q_z = \frac{(q/p)^z - 1}{(q/p)^M-1} $$
$$ q_z = \frac{(0.6/0.4)^1 - 1}{(0.6/0.4)^8-1} $$

$$ q_1 = 0.0203 $$

#### Experimental
```{r}
outcomes = c()
trials = 10000
for (i in (1:trials)){
  money = 1
  while (money > 0 && money < 8){
    if(rbinom(1,1,0.4)==1){
      money = money + 1
    }
    else {
      money = money - 1
    }
  }
  outcomes = append(outcomes, money)
}
hist(outcomes)
print(paste('Experimental probability of winning with timid strategy:',  (sum(outcomes)/8)/trials))
```

### Part B

(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).


#### Experimental
```{r}
outcomes = c()
trials = 10000
for (i in (1:trials)){
  money = 1
  while (money > 0 && money < 8){
    if(rbinom(1,1,0.4)==1){
      money = money + min(money, 8 - money)
    }
    else {
      money = money - min(money, 8 - money)
    }
  }
  outcomes = append(outcomes, money)
}
hist(outcomes)
print(paste('Experimental probability of winning with timid strategy:',  (sum(outcomes)/8)/trials))
```

### Part C

(c) Which strategy gives Smith the better chance of getting out of jail?

The bold strategy gives Smith the better chance of getting out of jail.