---
title: "Computational Mathematics - Assignment 11"
author: "Mary Anna Kivenson"
date: "November 9, 2019"
output:
  rmarkdown::html_document:
    
    theme: cosmo 
    highlight: zenburn
#    toc: true
#    toc_float: true
#    toc_depth: 4
---

```{r setup, include=FALSE}
library(ggfortify)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


# Regression - HW 11

## Cars - With Outliers {.tabset}

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

```{r}
data(cars)
head(cars, 6)
```


### Visualization

A scatter plot of stopping distance as a function of speed.

```{r}
ggplot(cars, aes(speed,dist)) + geom_point() + 
  ggtitle("Speed vs Stopping Distance") 
```

### Linear Model

```{r}
attach(cars)
lm <- lm(dist ~ speed)
lm
```

```{r}
eq = paste0("y = ", round(lm[1]$coefficients[2],3), "*x + ", round(lm[1]$coefficients[1],3))
ggplot(cars, aes(speed,dist)) + geom_point() + 
  geom_abline(intercept = lm[1]$coefficients[1], slope = lm[1]$coefficients[2]) +
  ggtitle(paste("Speed vs Stopping Distance:",eq)) 
```


### Model Quality

* This model has an `intercent` of 17.5791 and a `slope` of 3.9324, so for every additional unit of speed, the stopping distance increases by 3.9324. 
* The model has 48 degree of freedom, which is the number of samples minus the number of coefficents in the model (50 - 2).
* To determine the model quality, we look at the `Multiple R-squared` value. R-squared values closer to one indicate better model quality, so R^2 = 0.6511 indicates that this model is sufficient.

```{r}
summary(lm)
```

### Residual Analysis

The residuals do not have a trend, which indicates that a linear model fits the data well. However, there are more residual values below 0 than above 0, which suggests that the data may have outliers.

```{r}
cars$resid <- resid(lm)
ggplot(cars, aes(speed,resid)) + geom_point() + 
  geom_hline(yintercept=0, linetype="dashed", color = "blue") +
  ggtitle("Residual Plot") 
```


### Outliers

```{r}
dist_outlier <- boxplot.stats(cars$dist)$out
print(paste("There is a stopping distance outlier at",dist_outlier))
```
