---
title: "Computational Mathematics - Assignment 11"
author: "Mary Anna Kivenson"
date: "November 9, 2019"
output:
  rmarkdown::html_document:
    
    theme: cosmo 
    highlight: zenburn
#    toc: true
#    toc_float: true
#    toc_depth: 4
---

```{r setup, include=FALSE}
library(ggfortify)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


# Regression - Discussion 11

## Swiss Education vs Agriculture Regression {.tabset}

I am using the built-in R dataset of Swiss socioeconomic and fertility factors from 1888. I will focus on the relationship between education and agriculture. 

```{r}
data(swiss)
head(swiss, 6)
```


### Visualization

A scatter plot of agriculture as a function of education. We can take the log of education to fit the a linear regression.

```{r}
ggplot(swiss, aes(Education,Agriculture)) + geom_point() + 
  ggtitle("Education vs Agriculture") 

ggplot(swiss, aes(log(Education),Agriculture)) + geom_point() + 
  ggtitle("Log Education vs Agriculture") 
```

### Linear Model

Fit agriculture ~ log(education).

```{r}
attach(swiss)
lm <- lm(Agriculture ~ log(Education))
lm
```

```{r}
eq = paste0("y = ", round(lm[1]$coefficients[2],3), "*x + ", round(lm[1]$coefficients[1],3))
ggplot(swiss, aes(log(Education),Agriculture)) + geom_point() + 
  geom_abline(intercept = lm[1]$coefficients[1], slope = lm[1]$coefficients[2]) +
  ggtitle(paste("Log Education vs Agriculture:",eq)) 
```

### Conditions

* The Swiss towns which we used to fit the line are independent.

* The relationship between the log education and agriculture is linear.

* The residuals from the regression line are nearly normal.

* The variability of education/agriculture observations around the regression line is constant.


### Model Quality

* This model has an `intercent` of -19.348 and a `slope` of 91.277.

* To determine the model quality, we look at the `Multiple R-squared` value. R-squared values closer to one indicate better model quality, so R^2 =  0.4569 indicates that this model is not ideal.

```{r}
summary(lm)
```

### Residual Analysis

The residuals do not have a trend, which indicates that a linear model fits the data well.

```{r}
swiss$resid <- resid(lm)
ggplot(swiss, aes(log(Education),resid)) + geom_point() + 
  geom_hline(yintercept=0, linetype="dashed", color = "blue") +
  ggtitle("Residual Plot") 

ggplot(data=swiss, aes(swiss$resid)) + 
  geom_histogram(binwidth=5) +
  ggtitle("Residual Histogram") 
```
