---
title: "Matrix Multiplication"
author: "Mary Anna Kivenson"
date: "September 1, 2019"
output:
  rmarkdown::html_document:
    
    theme: simplex
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Main Example
Compute the product of the matrices below (Question C21, Page 190).

#### Question C21, Page 190

**Matrix A:**
```{r matrix_A, echo=FALSE}
A <- rbind(c(1,3,2),c(-1,2,1),c(0,1,0))
print(A)
```

**Matrix B:**
```{r matrix_B, echo=FALSE}
B <- rbind(c(4,1,2),c(1,0,1),c(3,1,5))
print(B)
```

**Product**

I created a custom function to generate the matrix product of any two matrices. This function iterates between rows of matrix A and columns of matrix B to find the dot product.

```{r product}
matrix_mult <- function(matrix_A, matrix_B){
  C = matrix(nrow = nrow(matrix_A), ncol = ncol(matrix_B))
  for (b in 1:ncol(matrix_B)){
    for (a in 1:nrow(matrix_A)){
      C[a,b] = sum(matrix_A[a,]*matrix_B[,b])
    }
  }
  return(C)
}

print(matrix_mult(A,B))
```

## Other Examples

I also applied this function to non-square matrices to see if it still works.


#### Question C20, Page 190
Compute the product of the matrices below.

```{r product2}
#matrix A
A <- rbind(c(2,5),c(-1,3),c(2,-2))
print(A)

#matrix B
B <- rbind(c(1,5,-3,4),c(2,0,2,-3))
print(B)

#matrix product
print(matrix_mult(A,B))
```



#### Question C23, Page 190
Compute the product of the matrices below.

```{r product3}
#matrix A
A <- rbind(c(3,1),c(2,4),c(6,5), c(1,2))
print(A)

#matrix B
B <- rbind(c(-3,1),c(4,2))
print(B)

#matrix product
print(matrix_mult(A,B))
```